---
title: "Exam 2-Jackson Freeman"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#clear environment
rm(list=ls(all=TRUE))
#set the working directory
setwd("C:/Users/Jack/Documents/Data Science 2020")
#import data as data frame
inequality_data<-import("inequality.xlsx")
```

1.) This data is cross-sectional-it is taken at one point in time (2015). Using the rio package and the HEAD function, we can show this concept.
```{r, include=True}
head(inequality_data)
```
2.) Subset for Gini of Denmark and Sweden

```{r gini subset for Den+Sweden, include=FALSE}
subset(inequality_data, c(country=="Sweden", country=="Denmark"), select=inequality_gini) 
```
3.) Subset for Brazil
```{r gini, include=FALSE}
subset(inequality_data, country=="Brazil", select=inequality_gini)
```
4.) Comparing these two countries, we can see that having a low inequality_gini score is better for a country.
5.) Here we will show the head of the data frame again.
```{r, include=TRUE}
head(inequality_data)
```
6.) Let's write a function for removing accents.
```{r removing accents, include=TRUE}
remove.accents<- function(s){
old1 <- "ú"
new1 <- "u"
s1 <- chartr(old1, new1, s)
}

#now we will run the function over the data set
inequality_data$country = remove.accents(inequality_data$country)
#Running head command to show removed accents
head(inequality_data)
```

7.) Sorting countries by lowest gini score.
```{r sorting data, include=FALSE}
inequality_data<- arrange(inequality_data,inequality_gini)

```
```{r, include=True}
head(inequality_data)
```

8.) Mean gini score. 
```{r mean gini score, include=TRUE}
mean(inequality_data$inequality_gini, na.rm=TRUE)
```
9.) Creating dummy variables and ifelse
```{r if else, include=FALSE}
#creating dummy variables 
#load tidyverse
library(tidyverse)
inequality_data <-inequality_data %>%
mutate(highinequality = ifelse(inequality_data$inequality_gini > 36.81375,
1,
0))

inequality_data <-
inequality_data %>%
mutate(lowinequality = ifelse(inequality_data$inequality_gini < 36.81375,
1,
0))
```
10.)Cross-tab of data.
```{r cross tab, include=TRUE}
library(doBy)
summaryBy( inequality_gini ~ highinequality, data=inequality_data, FUN=c(mean,length))
summaryBy( inequality_gini ~ lowinequality, data=inequality_data, FUN=c(mean,length))
```
11.)Creating a for loop for 3 actors.
```{r for loop, include=TRUE}
# Create an organization vector
orgs <- c(' World Bank', 'African Development Bank', 'Bill and Melinda Gates Foundation')
# Create the for statement
for ( i in orgs){
print(i)
}
```
12.) I am choosing the World Bank variable of "Poverty headcount ratio at $1.90 a day (2011 PPP) (% of population)" since it will show the percentage of population under the poverty line which can be used for inequality comparisons, especially with other variables like GDP in total.
13.) Let's import that data into r.
```{r import, include=TRUE}
#use WDI library
library(WDI)
# After Googling: https://data.worldbank.org/indicator/SI.POV.DDAY
population_data = WDI(country = "all",
                      indicator = c("SI.POV.DDAY"), # indicator from web
                      start = 2015, end = 2015, extra = FALSE, cache = NULL)
#let's now rename the variable
library(data.table)
setnames(population_data, "SI.POV.DDAY" , "population_poverty")

#Let's merge this data
#Remove NA's first
population_data <-
population_data %>%
dplyr::filter(!(country=="NA"))

library(tidyverse)
merged_data = left_join(x=inequality_data,
y=population_data,
by =c("country", "year"))
#iso2's don't match, let's find and eliminate them
library(tidyverse)
merged_data <-
merged_data %>%
mutate(iso2c_match = ifelse(iso2c.x == iso2c.y,
"yes",
"no"))
# list the instances in which the spelling doesn't match
subset(merged_data, iso2c_match =="no")
#drop iso2c.y and rename x
merged_data <-
merged_data %>%
select(-c("iso2c.y"))
  
setnames(population_data, "iso2c.x" , "iso2c code", skip_absent = TRUE)
#remove isomatch var
merged_data$iso2c_match=NULL

```
14.) Create a new data frame with no missing variables.
```{r omit, include=FALSE}
merged_df <- na.omit(merged_data, select=c("inequality_gini", "population in poverty"))
```
15.)Create new data frame over 30 gini
```{r omit, include=FALSE}
data_greater_30 <- merged_df %>%
dplyr::filter(inequality_gini > 30)

```
16.)Using data_greater_30, use to R to count how many countries have the sequence “ai”in their name.
```{r omit, include=FALSE}
grep("ai",data_greater_30)

```
16.) Use any command from the apply family to take the sum of inequality_gini in data_greater_30.
```{r omit, include=FALSE}

sum_data<-sapply(data_greater_30$inequality_gini, sum, na.omit=TRUE)
#print
sum_data
```
17.) Labelling Variables
```{r omit, include=FALSE}
# label the variables of merged_df
#install.packages("labelled")
library(labelled)
var_label(merged_df) <- list(`country` = "Country",
`year` = "year",
`lowinequality` = "Low Inequality Dummy Variable",
`highinequality` = "High Inequality Dummy Variable",
`population in poverty` = "Poverty Headcount as Ratio",
`inequality_gini` = "Inequality Gini Score",
`iso2c.x`= "ISO2 Country Code")
```
18.) Saving as Stata
```{r omit, include=FALSE}
# save the dataset in Stata format with the labels
library(rio)
export(merged_df, file = "final_data.dta")
#problem with export
setnames(merged_df, "iso2c.x" , "iso2c code", skip_absent = TRUE)
#works 
```

